Models

### 2021-09-17 analysis/05_model_metrics.Rmd
Add some explanatory comments

```{r}
library(tidyverse)
#library(moderndive)
library(lme4)
library(broom.mixed)
library(lmerTest)
```
#### Read in data, get final dataset for models
Get data and put it in long form 
```{r}
pins::board_register_github(repo = "karinorman/biodivTS_data", branch = "master")

@@

pins::pin(model_data, board = "github")
```
Add human footprint data
```{r}
# hfp_data <- pins::pin_get("hfp-data", board = "github")
# 
# model_data_hfp <- model_data %>% 
#   #dplyr::select(rarefyID, year) %>% 
#   #distinct() %>%
#   mutate(hfp_year = case_when(
#     year < 2001 ~ 1993,
#     TRUE ~ 2009
#   ), 
#   hfp_rarefyID = str_replace_all(rarefyID, pattern = "_bird", ""), 
#   hfp_rarefyID = str_replace_all(hfp_rarefyID, pattern = "_mamm", "")) %>%
#   left_join(hfp_data %>% dplyr::select(-STUDY_ID) %>% mutate(year = as.numeric(year)), by = c("hfp_year" = "year", "hfp_rarefyID" = "rarefyID")) %>%
#   dplyr::select(-hfp_year)

@@

Convergence problems!! See here:
https://joshua-nugent.github.io/allFit/
https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html
https://biologyforfun.wordpress.com/2018/04/09/help-i-have-convergence-warnings/

@@

```{r}
# SES_rand_slope <- model_data %>%
#   filter(metric %in%  c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_next")) %>%
#   group_nest(metric) %>%
#   mutate(#year_center = map(data, ~scale(.x$year)),
#          model = map(data, ~lmer(value ~ year_scaled + (year_scaled|study_id/rarefyID), 
#                                  data = .x)),
#            # null_model = map(data, ~lmer(value ~ (year_scaled|study_id/rarefyID), 
#            #                       data = .x)),
#            # test = map2(model, null_model, anova),
#            # test_res = map(test, tidy),
#            coef = map(model, tidy),
#            ) %>% 
#   select(metric, model, coef) %>%
#   unnest(cols = c(coef)) %>%
#   filter(term == "year_scaled") %>%
#   select(-c(effect, group, term, statistic))
# 
# #check that they all used the same number of observations
# map(SES_rand_slope$model, nobs)
# 
# #these two metrics are cranky and want their own year scaling to converge
# 
# #Jaccard Base
# jacc_base_mod <-  model_data %>%
#   filter(metric == ("Jaccard_base")) %>%
#   mutate(year_scaled = scale(year)) %>%
#   group_nest(metric) %>%
#   mutate(model = map(data, ~lmer(value ~ year_scaled + (year_scaled|study_id/rarefyID), 
#                                  data = .x)),
#            coef = map(model, tidy),
#            ) %>% 
#   select(metric, model, coef) %>%
#   unnest(cols = c(coef)) %>%
#   filter(term == "year_scaled") %>%
#   select(-c(effect, group, term, statistic))
# 
# # S model
# s_mod <- model_data %>%
#   filter(metric %in%  c("S")) %>%
#   #it wants its own scaling to converge for some reason
#   mutate(year_scaled = scale(year)) %>%
#   group_nest(metric) %>%
#   mutate(model = map(data, ~lmer(logvalue ~ year_scaled + (year_scaled|study_id/rarefyID), 
#                                  data = .x, REML = TRUE)),
#            coef = map(model, tidy)
#            ) %>% 
#   select(metric, model, coef) %>%
#   unnest(cols = c(coef)) %>%
#   filter(term == "year_scaled") %>%
#   select(-c(effect, group, term, statistic))
# 
# metric_model_table <- bind_rows(SES_rand_slope, s_mod, jacc_base_mod)
# 
# #save coefficients and models
# usethis::use_data(metric_model_table)
# pins::pin(metric_model_table, board = "github")
# ```
# Taxa model
# ```{r}
# taxa_rand_slope <- model_data %>%
#   filter(metric %in%  c("SES_FRic", "SES_FEve", "SES_FDiv")) %>%
#   group_nest(metric) %>%
#   mutate(#year_center = map(data, ~scale(.x$year)),
#          model = map(data, ~lmer(value ~ year_scaled + (year_scaled|study_id/rarefyID) + (year_scaled|taxa), 
#                                  data = .x)),
#            # null_model = map(data, ~lmer(value ~ (year_scaled|study_id/rarefyID), 
#            #                       data = .x)),
#            # test = map2(model, null_model, anova),
#            # test_res = map(test, tidy),
#            coef = map(model, tidy),
#            ) %>% 
#   select(metric, model, coef) %>%
#   unnest(cols = c(coef))
# 
# #Jaccard Base
# jacc_taxa_base_mod <-  model_data %>%
#   filter(metric == ("Jaccard_base")) %>%
#   mutate(year_scaled = scale(year)) %>%
#   group_nest(metric) %>%
#   mutate(model = map(data, ~lmer(value ~ year_scaled + (year_scaled|study_id/rarefyID) + (year_scaled|taxa), 
#                                  data = .x)),
#            coef = map(model, tidy),
#            ) %>% 
#   select(metric, model, coef) %>%
#   unnest(cols = c(coef)) %>%
#   filter(term == "year_scaled") %>%
#   select(-c(effect, group, term, statistic))
# 
#  s_taxa_mod <- model_data %>%
#   filter(metric %in%  c("S")) %>%
#   #it wants its own scaling to converge for some reason
#   mutate(year_scaled = scale(year)) %>%
#   group_nest(metric) %>%
#   mutate(model = map(data, ~lmer(logvalue ~ year_scaled + (year_scaled|study_id/rarefyID) + (year_scaled|taxa), 
#                                  data = .x, REML = TRUE)),
#            coef = map(model, tidy)
#            ) %>% 
#   select(metric, model, coef) %>%
#   unnest(cols = c(coef)) %>%
#   filter(term == "year_scaled") %>%
#   select(-c(effect, group, term, statistic))
# 
# fric_taxa_mod <- model_data %>%
#   filter(metric %in%  c("SES_FRic")) %>%
#   #it wants its own scaling to converge for some reason
#   mutate(year_scaled = scale(year)) %>%
#   group_nest(metric) %>%
#   mutate(model = map(data, ~lmer(logvalue ~ year_scaled + (year_scaled|study_id/rarefyID) + (year_scaled|taxa), 
#                                  data = .x, REML = TRUE)),
#            coef = map(model, tidy)
#            ) %>% 
#   select(metric, model, coef) %>%
#   unnest(cols = c(coef)) %>%
#   filter(term == "year_scaled") %>%
#   select(-c(effect, group, term, statistic))
# 
# taxa_model_table <- bind_rows(taxa_rand_slope, jacc_taxa_base_mod, s_taxa_mod, fric_taxa_mod)

@@

  year_scaled_coef <- tidy(model) %>%
    filter(term == "year_scaled")

  year_scaled_est <- pull(year_scaled_coef, estimate)

  year_scaled_var <- year_scaled_coef %>% 
    mutate(var = std.error*std.error) %>%
    pull(var)

  study_ests <- broom.mixed::tidy(model, effects="ran_vals") %>%
    filter(group == filter_name, term == "year_scaled") %>% 
    select(-c(effect, group, term)) %>%
    #rename({{ model_id }} := level, cond.std.error = std.error, cond.estimate = estimate) %>%
    rename(cond.std.error = std.error, cond.estimate = estimate) %>%
    mutate(estimate = cond.estimate + year_scaled_est,
      cond.var = cond.std.error*cond.std.error, 
           var = cond.var + year_scaled_var,
           std.error = sqrt(var),
           upr.ci = estimate + (2*std.error),
           lwr.ci = estimate - (2*std.error),
           sig = case_when(
             lwr.ci < 0 & upr.ci > 0 ~ FALSE,
             TRUE ~ TRUE
           ),
      metric = metric,
      model_id = model_id)
}

@@

```{r}
# #And how many were significant for all but Jaccard?
# study_slopes %>% 
#   filter(metric %in% c("SES_FRic", "SES_FEve", "SES_FDiv", "S")) %>% 
#   group_by(study_id) %>% filter(sum(sig) == 4) %>% pull(study_id) %>% 
#   n_distinct()
# 
# #Breakdown of relationship between functional richness and species richness
# s_sig <- study_slopes %>% filter(metric == "S", sig == TRUE) %>% pull(study_id) %>% unique()
# s_sig_not_fric <- study_slopes %>% 
#   filter(study_id %in% s_sig, metric == "FRic", sig == FALSE) %>% 
#   pull(study_id) %>%
#   unique()
# 
# both_sig <- study_slopes %>% 
#   filter(study_id %in% s_sig, metric == "FRic", sig == TRUE) %>%
#   pull(study_id)
# 
# both_pos
# fric_pos
# s_pos_fric_neg <- study_slopes %>%
#   filter(study_id %in% both_sig) %>% 
#   group_by(study_id) %>%
#   filter(sum(sign(estimate)) == 0)

>>>>

```{r}
library(tidyverse)
library(lme4)
library(broom.mixed)
library(lmerTest)
```
#### Read in data, get final dataset for models
Get data and put it in long form 
```{r}
pins::board_register_github(repo = "karinorman/biodivTS_data", branch = "master")

@@

pins::pin(model_data, board = "github")
```

Option to add human footprint data
```{r}
# hfp_data <- pins::pin_get("hfp-data", board = "github")
# 
# model_data_hfp <- model_data %>% 
#   #dplyr::select(rarefyID, year) %>% 
#   #distinct() %>%
#   mutate(hfp_year = case_when(
#     year < 2001 ~ 1993,
#     TRUE ~ 2009
#   ), 
#   hfp_rarefyID = str_replace_all(rarefyID, pattern = "_bird", ""), 
#   hfp_rarefyID = str_replace_all(hfp_rarefyID, pattern = "_mamm", "")) %>%
#   left_join(hfp_data %>% dplyr::select(-STUDY_ID) %>% mutate(year = as.numeric(year)), by = c("hfp_year" = "year", "hfp_rarefyID" = "rarefyID")) %>%
#   dplyr::select(-hfp_year)

@@

Convergence problems! See here:
https://joshua-nugent.github.io/allFit/
https://rstudio-pubs-static.s3.amazonaws.com/33653_57fc7b8e5d484c909b615d8633c01d51.html
https://biologyforfun.wordpress.com/2018/04/09/help-i-have-convergence-warnings/

@@

@@

  #get the statistics for the year variable
  year_scaled_coef <- tidy(model) %>%
    filter(term == "year_scaled")

  #get the year estimate
  year_scaled_est <- pull(year_scaled_coef, estimate)

  #get year variation
  year_scaled_var <- year_scaled_coef %>% 
    mutate(var = std.error*std.error) %>%
    pull(var)

  #get study level conditional estimates and add to over all estimates
  study_ests <- broom.mixed::tidy(model, effects="ran_vals") %>%
    filter(group == filter_name, term == "year_scaled") %>% 
    select(-c(effect, group, term)) %>%
    rename(cond.std.error = std.error, cond.estimate = estimate) %>%
    mutate(estimate = cond.estimate + year_scaled_est,
      cond.var = cond.std.error*cond.std.error, 
           var = cond.var + year_scaled_var,
           std.error = sqrt(var),
           upr.ci = estimate + (2*std.error),
           lwr.ci = estimate - (2*std.error),
           sig = case_when(
             lwr.ci < 0 & upr.ci > 0 ~ FALSE,
             TRUE ~ TRUE
           ),
      metric = metric,
      model_id = model_id)
}

@@
